Real-space computation of $E$/$B$-mode maps I: Formalism and Compact Kernels

### 2017-11-02 articles/formalism/sections/generalizedeb.tex
updated color scheme for line plots

While defining these generalized operators, it seems more natural to choose the real space function $G(\beta)$ as compared to choosing the multipole function $g_{\ell}$. Using the orthogonality property of associated Legendre polynomials it can be shown that the harmonic function $g_{\ell}$ is given by the following integral over the radial function $G(\beta)$,
%
\beq
g_{\ell} = 2 \pi \sqrt{\frac{(\ell-2)!}{(\ell+2)!}} \int _{0}^{\pi} G(\beta) P_{\ell}^{2}(\cos{\beta}) d\cos{\beta} \,. \label{eq:gb2bl}
\eeq
%
Here it is important to note that the radial function $G(\beta)$ has to be chosen such that it vanishes at $\beta=0$ and $\beta=\pi$. 
%One way to understand this is that the associated Legendre polynomials $P_{\ell}^2 \propto \sin^2{\beta}$ vanish at these values of the abscissa and hence cannot be used to describe functions which don't have this property.  Another way to understand this requirements is that at these locations the coordinate dependence of the Stokes parameters cannot be integrated out, since the azimuthal angle is ill defined and hence the convolution kernel needs to have vanishing contribution from these locations. \revisit{The normalization of these functions is not critically important, since any choice defines a convention. It is important to ensure consistency with the convention once a choice has been made.} \comment{find a better location for this}
Note that in contrast to the radial function $G(\beta)$ an instrumental beam function appropriately normalized has the property $B(\beta) \rightarrow 1$ as $\beta \rightarrow 0$. We clarify that the $B(\beta)$ refers to the effective beam acting to smooth the scalar $E$ \& $B$ mode maps. A circularly symmetric beam $B(\beta)$ defined at the pole can be expressed in the Legendre polynomial $P_{\ell}^0$ basis as follows,
%
\beq
B(\beta) = \sum_{\ell=0}^{\ell_{\rm max}} \frac{2 \ell+1}{4 \pi} b_{\ell} P_{\ell}^{0} (\cos{\beta})\,,
\eeq
%
where $b_{\ell}$ denote the coefficients of expansion.
%
\begin{figure}[!t] 
\centering
\subfigure[\label{fig:bl_gbeta}]{\includegraphics[width=0.32\columnwidth]{Gbeta_for_different_gl_lmax1536.pdf}}
\subfigure[\label{fig:glbl}]{\includegraphics[width=0.32\columnwidth]{gl_for_different_gbeta_lmax1536.pdf}}
\subfigure[\label{fig:gl_bbeta}]{\includegraphics[width=0.32\columnwidth]{Bbeta_for_gl_lmax1536.pdf}}
\caption{\textit{Left:} The vertical dashed gray line depicts the approximate pixel size $\Delta_{\rm pix} = \sqrt{\frac{4 \pi}{N_{\rm pix}}}$ of a Nside=512 Healpix map. The green line depicts the default radial kernel $f(\beta)$ defined in \eq{eq:qu2eb_gen_kernel}. The blue and orange lines depict the modified radial function resulting the beam harmonics $b_{\ell}$ corresponding to Gaussian beams with fwhm=15 \& 12 arcminutes respectively. The red curve depicts an example modified radial function: $G(\beta)=\mathcal{N} \beta^n \exp{\left[ -\left( \frac{\beta-\beta0}{\sqrt{2} \sigma} \right)^s \right]}$ with parameters set to the following values $[n=1;\, \beta_0=0 ;\, \sigma = 2\Delta_{\rm pix} ;\, s=1.5]$. The black dashed curve depicts the band limited reconstruction of the modified radial function $G(\beta)$. We intentionally have plotted $\beta G(\beta)$ to clearly depict the high $\beta$ behavior of these functions. \textit{Middle: } This figure depicts the harmonic representation of the respective radial functions as indicated by the legend. The dashed curves of the corresponding color depict the inverse of the harmonic functions. \textit{Right:} This figure depicts the beam function $B(\beta)$ evaluated from interpreting the respective harmonic functions as those corresponding to an instrument beam.}
\label{fig:example_gbeta}
\end{figure}
%

>>>>

While defining these generalized operators, it seems more natural to choose the real space function $G(\beta)$ as compared to choosing the multipole function $g_{\ell}$. Using the orthogonality property of associated Legendre polynomials it can be shown that the harmonic function $g_{\ell}$ is given by the following integral over the radial function $G(\beta)$,
%
\beq
g_{\ell} = 2 \pi \sqrt{\frac{(\ell-2)!}{(\ell+2)!}} \int _{0}^{\pi} G(\beta) P_{\ell}^{2}(\cos{\beta}) d\cos{\beta} \,. \label{eq:gb2bl}
\eeq
%
Here it is important to note that the radial function $G(\beta)$ has to be chosen such that it vanishes at $\beta=0$ and $\beta=\pi$. 
%One way to understand this is that the associated Legendre polynomials $P_{\ell}^2 \propto \sin^2{\beta}$ vanish at these values of the abscissa and hence cannot be used to describe functions which don't have this property.  Another way to understand this requirements is that at these locations the coordinate dependence of the Stokes parameters cannot be integrated out, since the azimuthal angle is ill defined and hence the convolution kernel needs to have vanishing contribution from these locations. \revisit{The normalization of these functions is not critically important, since any choice defines a convention. It is important to ensure consistency with the convention once a choice has been made.} \comment{find a better location for this}
Note that in contrast to the radial function $G(\beta)$ an instrumental beam function appropriately normalized has the property $B(\beta) \rightarrow 1$ as $\beta \rightarrow 0$. We clarify that the $B(\beta)$ refers to the effective beam acting to smooth the scalar $E$ \& $B$ mode maps. A circularly symmetric beam $B(\beta)$ defined at the pole can be expressed in the Legendre polynomial $P_{\ell}^0$ basis as follows,
%
\beq
B(\beta) = \sum_{\ell=0}^{\ell_{\rm max}} \frac{2 \ell+1}{4 \pi} b_{\ell} P_{\ell}^{0} (\cos{\beta})\,,
\eeq
%
where $b_{\ell}$ denote the coefficients of expansion.
%
\begin{figure}[!t] 
\centering
\subfigure[\label{fig:glbl}]{\includegraphics[width=0.48\columnwidth]{gl_for_different_gbeta_lmax1536.pdf}}\\
\subfigure[\label{fig:bl_gbeta}]{\includegraphics[width=0.48\columnwidth]{Gbeta_for_different_gl_lmax1536.pdf}}
\subfigure[\label{fig:gl_bbeta}]{\includegraphics[width=0.48\columnwidth]{Bbeta_for_gl_lmax1536.pdf}}
\caption{\textit{Left:} The vertical dashed gray line depicts the approximate pixel size $\Delta_{\rm pix} = \sqrt{\frac{4 \pi}{N_{\rm pix}}}$ of a Nside=512 Healpix map. The green line depicts the default radial kernel $f(\beta)$ defined in \eq{eq:qu2eb_gen_kernel}. The blue and orange lines depict the modified radial function resulting the beam harmonics $b_{\ell}$ corresponding to Gaussian beams with fwhm=15 \& 12 arcminutes respectively. The red curve depicts an example modified radial function: $G(\beta)=\mathcal{N} \beta^n \exp{\left[ -\left( \frac{\beta-\beta0}{\sqrt{2} \sigma} \right)^s \right]}$ with parameters set to the following values $[n=1;\, \beta_0=0 ;\, \sigma = 2\Delta_{\rm pix} ;\, s=1.5]$. The black dashed curve depicts the band limited reconstruction of the modified radial function $G(\beta)$. We intentionally have plotted $\beta G(\beta)$ to clearly depict the high $\beta$ behavior of these functions. \textit{Middle: } This figure depicts the harmonic representation of the respective radial functions as indicated by the legend. The dashed curves of the corresponding color depict the inverse of the harmonic functions. \textit{Right:} This figure depicts the beam function $B(\beta)$ evaluated from interpreting the respective harmonic functions as those corresponding to an instrument beam.}
\label{fig:example_gbeta}
\end{figure}
%
